---
title: "Biostat M280 Homework 2"
subtitle: Due Feb 16 @ 11:59PM
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = TRUE)
```

## Q1

Read [Chapter 7](http://r4ds.had.co.nz/exploratory-data-analysis.html) (Exploratory Data Analysis) of _R for Data Science_ and do exercises 7.3.4, 7.4.1, 7.5.1.1, 7.5.2.1, and 7.5.3.1.

    I read Chapter 7 and did exercises 7.3.4, 7.4.1, 7.5.1.1, 7.5.2.1, and 7.5.3.1.

7.3.4 
1. Explore the distribution of each of the x, y, and z variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.

    The distribution of the x, y, and z variables in diamonds are right skewed, indicating that the sizes of most diamonds are small and only a few are large. 
    
    The distribution of the x is very similar to that of y. In my opinion, x should be the length and y should be the width since they are close but the value of x is larger (more to the right) than that of y. And z should be the depth.
    ```{r, echo=TRUE, eval=TRUE}
    if(!"tidyverse" %in% rownames(installed.packages()))
        install.packages("tidyverse", repos = "http://cran.rstudio.com/")
    library(tidyverse)
    
    diamonds %>%
      select(x, y, z) %>%
      gather(variable, value) %>%
      ggplot(aes(x = value)) +
      geom_density() +
      facet_grid(variable ~ .)
    ```

2. Explore the distribution of price. Do you discover anything unusual or surprising? (Hint: Carefully think about the binwidth and make sure you try a wide range of values.)

    In general, as the price of the diamonds went higher, the number of diamonds went lower.
    
    The number of diamonds is 0 when the price is 1500. 
    ```{r, echo=TRUE, eval=TRUE}
    ggplot(data = diamonds) +
      geom_histogram(aes(x = price), binwidth = 100)
    
    ggplot(data = diamonds) + 
      geom_histogram(aes(x = price), binwidth = 10) +
      xlim(c(0, 2500))
    ```

3. How many diamonds are 0.99 carat? How many are 1 carat? What do you think is the cause of the difference?

4. Compare and contrast coord_cartesian() vs xlim() or ylim() when zooming in on a histogram. What happens if you leave binwidth unset? What happens if you try and zoom so only half a bar shows?




## Q2 (optional)

Read [Chapter 23](http://r4ds.had.co.nz/model-basics.html) (Model Basics) and [Chapter 24](http://r4ds.had.co.nz/model-building.html) (Model Building) of _R for Data Science_ and do exercises 24.2.3 and 24.3.5.

## Q3

Redo HW1 Q2 using tidyverse.

## Q4 (optional)

Redo HW1 Q3 on Hoffman2, except now we want to submit each `runSum.R` job to a different node in the cluster.