---
title: "Biostat M280 Homework 3"
subtitle: Due Mar 2 @ 11:59PM
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1 LA City Employee Payroll

The `/home/m280-data/la_payroll/LA_City_Employee_Payroll.csv` file on teaching server contains payroll information of LA City employees in years 2013-2017. It was downloaded from [LA City Controller's Office](https://controllerdata.lacity.org/Payroll/City-Employee-Payroll/pazn-qyym). Make a Shiny app to facilitate exploratory data analysis. 

1. For efficiency of the Shiny app, you should first pre-process, pare down, tidy, and save the data, e.g., as a compressed RDS file, to be used in the app.

    I pre-process, pare down, tidy, and save the data as a compressed RDS file. 
    ```{r}
    if(!"rio" %in% rownames(installed.packages()))
            install.packages("rio", repos = "http://cran.rstudio.com/")
    library(rio)
    library(tidyverse)
    convert("/home/m280-data/la_payroll/LA_City_Employee_Payroll.csv",
            "payroll_origin.rds")
    ```

2. **Total payroll by LA City**. Visualize the total LA City payroll of each year, with breakdown into base pay, overtime pay, and other pay.

```{r}
range(payroll$Year)
```



3. **Who earned most?** Visualize the payroll information (total payment with breakdown into base pay, overtime pay, and other pay, Department, Job Title) of the top $n$ highest paid LA City employees in a specific year. User specifies $n$ (default 10) and year (default 2017).

4. **Which departments earn most?** Visualize the mean or median payroll, with breakdown into base pay, overtime pay, and other pay, of top $n$ earning departments. User specifies $n$ (default 5), year (default 2017), and method (mean or median, default median).

5. **Which departments cost most?** Visualize the total payroll, with breakdown into base pay, overtime pay, and other pay, of top $n$ expensive departments. User specifies $n$ (default 5) and year (default 2017).

6. Visualize any other information you are interested in.

7. Publish your Shiny app to <https://www.shinyapps.io> and share the link.

## Q2 LA City Parking War

The SQLite database `/home/m280-data/la_parking/LA_Parking_Citations.sqlite` on teaching server contains information about parking tickets in LA City. It was downloaded from [LA Open Data Portal](https://data.lacity.org/A-Well-Run-City/Parking-Citations/wjz9-h9np). Connect to the database and answer following questions using plots and summary statistics. In this exercise, you are **not** allowed to load whole data into memory. Use the _transform in database, plot in R_ strategy.

    ```{r}
    if(!"DBI" %in% rownames(installed.packages()))
            install.packages("DBI", repos = "http://cran.rstudio.com/")
    suppressMessages(library(DBI))
    if(!"RSQLite" %in% rownames(installed.packages()))
            install.packages("RSQLite", repos = "http://cran.rstudio.com/")
    suppressMessages(library(RSQLite))
    if(!"dplyr" %in% rownames(installed.packages()))
            install.packages("dplyr", repos = "http://cran.rstudio.com/")
    suppressMessages(library(dplyr))
    if(!"lubridate" %in% rownames(installed.packages()))
            install.packages("lubridate", repos = "http://cran.rstudio.com/")
    suppressMessages(library(lubridate))
    if(!"tidyverse" %in% rownames(installed.packages()))
            install.packages("tidyversee", repos = "http://cran.rstudio.com/")
    suppressMessages(library(tidyverse))
    parking <- dbConnect(RSQLite::SQLite(), 
      dbname = "/home/m280-data/la_parking/LA_Parking_Citations_Extra.sqlite")
    dbListTables(parking)
    latix_sql <- dplyr::tbl(parking, "latix")
    latix_sql
    ```

1. How many tickets are in this data set? Which time period do these tickets span? Which years have most data?

    There are 4044338	tickets in the data set. 
    ```{sql connection=parking}
    SELECT COUNT(DISTINCT Ticket_number) FROM latix
    ```
    
    The time period of these tickets is from 2010/04/27 21:40 to 2017/12/30 01:41.
    ```{sql connection=parking}
    SELECT 
    MAX(Issue_Year || '/' || SUBSTR('0' || Issue_Month, -2 ) 
    || '/' || SUBSTR('0' || Issue_Day, -2) || ' ' 
    || SUBSTR('0' || Issue_Hour, -2) || ':' 
    || SUBSTR('0' || Issue_Minute, -2)) as 'Max', 
    MIN(Issue_Year || '/' || SUBSTR('0' || Issue_Month, -2 ) 
    || '/' || SUBSTR('0' || Issue_Day, -2) || ' ' 
    || SUBSTR('0' || Issue_Hour, -2) || ':' 
    || SUBSTR('0' || Issue_Minute, -2)) as 'Min'
    FROM latix
    ```
    
    The year 2015 has the most data.
    ```{r}
    latix_sql %>%
      filter(!is.na(Issue_Year)) %>%
      group_by(Issue_Year) %>%
      count() %>%
      collect() %>%
      ggplot(latix_sql, mapping = aes(x = Issue_Year, y = n)) +
      scale_x_continuous(breaks = c(2010:2017)) +
      geom_col()
    ```
    
2. When (which hour, weekday, month day, and month) are you most likely to get a ticket and when are you least likely to get a ticket?

    You are most likely to get a ticket at hour 12. You are least likely to get a ticket at hour 5.
    ```{r}
    latix_sql %>%
      filter(!is.na(Issue_Hour)) %>%
      group_by(Issue_Hour) %>%
      count() %>%
      collect() %>%
      ggplot(latix_sql, mapping = aes(x = Issue_Hour, y = n)) +
          geom_col()
    ```
    
    You are most likely to get a ticket on the 22nd. You are least likely to get a ticket on the 31st.
    ```{r}
    latix_sql %>%
      filter(!is.na(Issue_Wday)) %>%
      group_by(Issue_Wday) %>%
      count() %>%
      collect() %>%
      ggplot(latix_sql, mapping = aes(x = Issue_Wday, y = n)) +
          geom_col()
    ```

    You are most likely to get a ticket at Tuesday. You are least likely to get a ticket on Saturday.
    ```{r}
    latix_sql %>%
      filter(!is.na(Issue_Day)) %>%
      group_by(Issue_Day) %>%
      count() %>%
      collect() %>%
      ggplot(latix_sql, mapping = aes(x = Issue_Day, y = n)) +
      scale_x_continuous(breaks = c(1:31)) +
          geom_col()
    ```

    You are most likely to get a ticket at March. You are least likely to get a ticket at December.
    ```{r}
    latix_sql %>%
      filter(!is.na(Issue_Month)) %>%
      group_by(Issue_Month) %>%
      count() %>%
      collect() %>%
      ggplot(latix_sql, mapping = aes(x = Issue_Month, y = n)) +
      scale_x_continuous(breaks = c(1:12)) +
          geom_col()
    ```

3. Which car makes received most citations?

    The toyota received most citations. 
    ```{r}
    latix_sql %>%
      filter(!is.na(Make)) %>%
      group_by(Make) %>%
      summarise(num = n()) %>%
      collect() %>%
      top_n(5) %>%
      ggplot(latix_sql, mapping = aes(x = Make, y = num)) +
          geom_col()
    ```

4. How many different colors of cars were ticketed? Which color attracted most tickets?

    There are 65 different colors of cars were ticketed.
    ```{sql connection=parking}
    SELECT COUNT(DISTINCT Color) FROM latix
    ```

    The color black attracked most tickets. 
    ```{r}
    latix_sql %>%
      filter(!is.na(Color)) %>%
      group_by(Color) %>%
      summarise(num = n()) %>%
      collect() %>%
      top_n(5) %>%
      ggplot(latix_sql, mapping = aes(x = Color, y = num)) +
          geom_col()
    ```

5. What are the most common ticket types?

    NO PARK/STREET CLEAN, METER EXP, and PREFERENTIAL PARKING are the most common ticket types. 
    ```{r}
    latix_sql %>%
      filter(!is.na(Violation_Description)) %>%
      group_by(Violation_Description) %>%
      summarise(num = n()) %>%
      collect() %>%
      top_n(3) %>%
      ggplot(latix_sql, mapping = aes(x = Violation_Description, y = num)) +
          geom_col()
    ```
    
6. How much money was collected on parking tickets in 2015 and 2016?

    151006794 dollors were collected on parking tickets in 2015. 123236136 dollors were collected for 2016. 
    ```{sql connection=parking}
    SELECT 
    Issue_Year, SUM(Fine_amount)
    FROM latix
    GROUP BY Issue_Year
    ORDER BY SUM(Fine_amount) DESC
    ```

7. Visualize any other information you are interested in.

    I am interested in which states of the plate attracted most tickets.
    ```{r}
    latix_sql %>%
      filter(!is.na(RP_State_Plate)) %>%
      group_by(RP_State_Plate) %>%
      summarise(num = n()) %>%
      collect() %>%
      top_n(5) %>%
      ggplot(latix_sql, mapping = aes(x = RP_State_Plate, y = num)) +
          geom_col()
    ```

    ```{r}
    dbDisconnect(parking)
    ```


